---
title: "Exploration of the available data sources"
author: "Lukas Dargel"
date: "2021-03-04"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  autodep = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{r setup}
library("data.table")
library("dplyr")
library("here")
library("rrMD")
library("sf")
library("skimr")
```


```{r dataLoad}
source("R/helpers_data-import.R" %>% here())
load_raw_data <- get_raw_data_loaders()
```

# What information do we need?

The objective of the project is to create a gravity model of retail sales.
This models the commercial flows from the point were clients live to the stores of customers.

+ We need information on the origins of the customers which is on IRIS level.
+ We need information on the destinations which are the shops.
+ We need informations on pairs of origins and destinations (distance and amounts of flows).

To collect this information we have 4 internal sources of geomarketing data.  
Additional information has to be found on open sources which are INSEEE census data.

We consider the following data sources.
